Draft

### 2017-02-21 Draft/sections/misc-math.tex
Add the Zassenhaus formula for exponential of operators.
[major] Notes about a trick to calculate Jacobi matrix (det(A)=1+tr(A)).

>>>>

\subsection{Matrix Operations}
    \subsubsection{Matrix Exponentials}
    \label{sec:Matrix-Exponentials}
    I found these formulae from Wikipedia page:
    \href{https://en.wikipedia.org/wiki/Baker%E2%80%93Campbell%E2%80%93Hausdorff_formula}{Baker–Campbell–Hausdorff formula}.
    \paragraph{The Zassenhaus formula}
    \begin{align}
        e^{t(X+Y)} &= e^{tX}~ e^{tY} ~e^{-\frac{t^2}{2} [X,Y]} ~
        e^{\frac{t^3}{6}(2[Y,[X,Y]]+ [X,[X,Y]] )} \nonumber\\
        & \times
          e^{\frac{-t^4}{24}([[[X,Y],X],X] + 3[[[X,Y],X],Y] + 3[[[X,Y],Y],Y]) }
        \times\cdots
    \end{align}

    \paragraph{Logarithm of Matrix}
    The logarithm of  a matrix $A$ is defined when there is a matrix $B$ such that
    $e^B=A$, then $\ln(A)=B$.

    Using this, we have a formula to calculate the determinant of a some special
    matrix.

    Let $A = I + B$, and suppose $\det(A)\neq
    0$. Then, we have:
    \begin{equation}
        \det(A) = \exp(\tr(\ln(A)))
    \end{equation}
    Since $\exp{\sum_i \ln(\lambda_i)} = \Pi_i \exp(\ln(\lambda_i))=\Pi_i
    \lambda_i=\det(A)$.

    Next, we recap several series:
    \begin{align*}
        \ln(1+x) &= \sum_{n=1}^{\infty} \frac{(-1)^{n+1}}{n}x^n \\
        \exp(x) &= \sum_{n=0}^{\infty} \frac{x^n}{n!}
    \end{align*}
    Then we start the calculation:
    \begin{align*}
        \ln(I+B) &= \sum_{n=1}^{\infty} \frac{(-1)^{n+1}}{n}B^n \\
        \tr(\ln(I+B)) &= \sum_{n=1}^{\infty} \frac{(-1)^{n+1}}{n}\tr(B^n) \\
        \det(A) &= \exp(\tr(\ln(I+B))) 
            = \exp(\sum_{n=1}^{\infty} \frac{(-1)^{n+1}}{n}\tr(B^n)) \\
            &\approx 1+ \sum_{n=1}^{\infty} \frac{(-1)^{n+1}}{n}\tr(B^n) \\
            &\approx 1+ \tr(B)
    \end{align*}

    Therefore, we have, for appropriate matrix $B$:
    \begin{thm}
    \begin{equation}
        \det(A) \equiv \det(I+B) \approx 1+\tr(B)
    \end{equation}
    \end{thm}

    A typical situation is that $A$ is the transformation matrix:
    \begin{equation}
        A^\mu_\nu = \frac{\partial x'_\mu}{\partial x_\nu} 
        = \delta^\mu_\nu + \text{(first order matrix $B$)}^\mu_\nu
    \end{equation}
